### 2010-04-23 papers/wmt10/scratch.tex
Added a clean bibliography
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1610 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

%
% File acl2010.tex
%
% Contact  jshin@csie.ncnu.edu.tw or pkoehn@inf.ed.ac.uk
%%
%% Based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2010}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage[colorlinks]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
%\setlength\titlebox{6.5cm}    % You can expand the title box if you
% really have to

\title{Joshua: An Open Source Toolkit for Parsing-based Machine Translation}

\author{
Zhifei Li,\,\,\,
 Lane Schwartz,$^\star$\,\,\,
Chris Callison-Burch,\,\,\, %add a footnote to mention that CCB is the project leader
Chris Dyer,$^\dagger$\,\,\,
Juri Ganitkevitch,\,\,\, \\
{\bf Sanjeev Khudanpur,\,\,\, 
Wren N.\,G.\,Thornton,\,\,\,
Jonathan Weese\,\,
{\textnormal{and}}\,\,\,Omar F. Zaidan}\\
Center for Language and Speech Processing, Johns Hopkins University, Baltimore, MD\\
$\dagger$ Computational Linguistics and Information Processing Lab, University of Maryland, College Park, MD\\
$\star$ Natural Language Processing Lab, University of Minnesota, Minneapolis, MN }

\date{}

\begin{document}
\maketitle

\begin{abstract}
 This paper describes Joshua, an open source hierarchical phrase-based statistical machine translation system. Results for the WMT 2010 shared task are reported for six language pairs.
\end{abstract}

\listoftodos

\section{Grammar Extraction}

\todo{Add info about grammar extraction; MT Marathon paper and speedups}

\section{Results}

\todo{Describe results}

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l||c|c|c|}
\hline
\bf Source & \bf Target & \bf BLEU & \bf BLEU- & \bf TER \\
& & & \bf cased & \\
\hline
German & English & 21.3 & 19.5 & 0.660 \\ \hline
English & German & 15.2  & 14.6  & 0.738 \\ \hline
French & English & 27.7 & 26.4 & 0.614 \\ \hline
English & French & 23.8 & 22.8 & 0.681 \\ \hline
Spanish & English & 29.0 & 27.6 & 0.595 \\ \hline
English & Spanish & 28.1 & 26.5 & 0.596  \\ \hline
\end{tabular}
\end{center}
\caption{\label{scores} Automatic metric scores for the test set newstest2010 }
\end{table}

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l||c|}
\hline
\bf Source & \bf Target & \bf BLEU  \\
\hline
German & English & 18.19  \\ \hline
English & German & 13.57 \\ \hline
French & English & 26.41 \\ \hline
English & French & 25.28 \\ \hline
Spanish & English & 25.28 \\ \hline
English & Spanish & 24.02  \\ \hline
\end{tabular}
\end{center}
\caption{\label{devtest-scores} Automatic metric scores for the development test set newstest2009}
\end{table}

\end{document}